---
title: "NY Shooting"
author: 
date: "2023-03-15"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Project Overview
This report analyses the shooting cases in NY in the year 2022. Using visualizations, the primary goal is to observe and understand patterns in the data. 

## Importing data
The data use in this analysis can be found on the link <https://data.cityofnewyork.us/api/views/5ucz-vwe8/rows.csv?accessType=DOWNLOAD>.
After importing the data to RStudio, the following columns were removed: Longitude,New.Georeferenced.Column,X_COORD_CD,Y_COORD_CD, Latitude, LOCATION_DESC.

```{r dowlond, include=FALSE}
shooting_his <- read.csv("https://data.cityofnewyork.us/api/views/5ucz-vwe8/rows.csv?accessType=DOWNLOAD")
library(tidyverse)
library(lubridate)
```


```{r clean, include=FALSE}
shooting_his <- shooting_his %>%
  select(-c(Longitude,New.Georeferenced.Column,X_COORD_CD,Y_COORD_CD, Latitude, LOCATION_DESC,))%>%
  mutate(OCCUR_DATE = mdy(OCCUR_DATE))%>%
  mutate(OCCUR_TIME = hms(OCCUR_TIME))
```
After that, a first summary of the data was run:
```{r summary, echo=FALSE}
summary(shooting_his)
```

## Ploting

Firstly, is important to take a look at the profile of the victims and the perpetrator, keeping in mind that the second one was not always identifiable, there for the data is not complete.
```{r profile, echo=FALSE}
ggplot(shooting_his,aes(VIC_SEX, fill = VIC_RACE)) + geom_bar(position = "dodge")
ggplot(shooting_his,aes(PERP_SEX, fill = PERP_RACE)) + geom_bar(position = "dodge")
ggplot(shooting_his,aes(VIC_AGE_GROUP)) + geom_bar(fill="blue")
ggplot(shooting_his,aes(PERP_AGE_GROUP)) + geom_bar(fill="blue")
```

Is interesting to observe that the Victims and the perpetrator have a similar profile. Both are majority black males between the ages of 25-44. 

Now, looking into the crime itself we can obverse that most of them occur in Brooklyn and in the Bronx, with their pick in July. However, there isn't a direct relationship between the location of the shootings and the date.

```{r crime, echo=FALSE}
ggplot(shooting_his,aes(BORO)) + geom_bar(fill="blue")
ggplot(shooting_his,aes(OCCUR_DATE)) + geom_freqpoly() 
ggplot(shooting_his,aes(OCCUR_DATE,BORO)) + geom_jitter(height = 2, width = 2)
```

Also, observing the murder Flags, most shootings didn't result in death.
```{r unlive, echo=FALSE}
ggplot(shooting_his,aes(STATISTICAL_MURDER_FLAG)) + geom_bar(fill="blue")
```

## Avoiding Bias
As observed in the data, most of the people involved in the shooting were Black males. This is in the expected range, since is a pattern then repeated itself in violence statistics. A possible justification is that they represent the majority of the population. However, after consulting the 2021 census(2022 was not available), in the state of NY Black males only represent 6.8% of the population. Data can be founded at <https://data.census.gov/mdat/#/search?ds=ACSPUMS1Y2021&cv=SEX&rv=RACBLK&wt=PWGTP>. 
## Suggestions
After observing the data is possible to conclude that is a similar social profile for these crimes, with this information the city should invest in programs that help young black men, such as sports centers or programming classes. Also, the use of geoanalyses should be taken into consideration when patrols are being assigned. Lastly, is important to keep in mind the months with high violence risk, suck as July, scheduling patrol time off around that. 
